A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-04 tex/codatadp.tex
copied stuff into Sect. 2

>>>>

    semantically, $k$ types are mutually (co)recursive if their dependency graph
    (x -> y if x is the type of an argument to one of y's constructors) is
    strongly connected

    (co)datatypes can be declared together as mutually (co)recursive even if
    they are not actually (co)recursive

    the semantic notion is more fruitful and the one we will focus on in the
    sequel

    Also, nothing forbids specifications with no (co)recursion, e.g.

      datatype enat = Nat(nat) | Infty
      codatatype complex = Complex(re: real, im: real)

    In this case, it makes no difference whether the type is introduced as a
    datatype or as a codatatype. Without loss of generality, we consider that
    these types are datatypes instead of codatatypes.

Degenerate Codatatypes

    Consider

      codatatype a = A(a)

    This is a finite codatatype, with a single value, which is infinite:
    $\mu a. A(a)$

    We need to detect those finite codatatypes. They are easy to
    characterize semantically.

    Lemma D.1. The only possible finite cardinality for the domain interpreting
    a (corecursive) codatatype is 1.

    Proof.
      Assume you have at least 2 different values of a codatatype.
      They differ at one point.
      * 
    Qed

    An immediate consequence by the distinctness axiom is that these types have
    necessarly one constructor, and that all arguments are of types of
    cardinality 1.

    Easy case: $k$ mutually recursive codatatypes, each equipped with one
    constructor, all of which take each other in arguments in various
    combinations, with potential other types that are known to have cardinality
    1. e.g.

      codatatype b = B b c b
             and c = C c b c

    Medium case: like above, but some of the arguments are other types not
    participating in the corecursion that are known to have cardinality 1, e.g.

      codatatype d = D unit d d unit

    where "unit" is a datatype with the single constructor "Unity : unit".

    Though case: like above, but with any other types where it's not clear
    whether it could have cardinality one or not.

      type e
      codatatype f = F e f f unit
